Atomify - a live LAMMPS visualizer

### 2017-05-07 ms.tex
Added gui figure
Changed key

\documentclass[aps,pre,twocolumn,letterpaper,floatfix]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\usepackage{todo}
\begin{document}

@@

\section{Introduction}
%
Molecular dynamics (MD) is a common technique for simulating the motion of atoms
and molecules in a broad range of fields.
% TODO does MD also include larger particles?
% TODO which fields?
Over the years, increasingly sophisticated methods have been developed.
% TODO what methods?
With the large number of force fields, advanced techniques and GPU acceleration,
implementing a good molecular dynamics simulation is a challenging task.
There are many software packages for molecular dynamics modelling
available
\footnote{See \url{https://en.wikipedia.org/wiki/Comparison_of_software_for_molecular_mechanics_modeling}.}
Some packages, such as LAMMPS\cite{Plimpton1995Fast}, GROMACS\cite{Pronk2013}
and NAMD\cite{Phillips2005Scalable} have been developed for more than two decades 
and contain more than one million lines of code each.

The work in this paper is focused on LAMMPS.
Although the software is well-documented and enables the user to perform
advanced simulations, it can be demanding to learn how to use it well.
A LAMMPS simulation is performed by a series of commands which are executed one by
one.
Physical quantities such as energy, temperature and stress can be measured by
adding \textit{computes}.
The time evolution of these quantities can in turn be saved to file for plotting
and further analysis.
Particle trajectories can be saved to a file and visualized by tools such as
VMD\cite{Humphrey1996Vmd} or OVITO\cite{Stukowski2009Visualization}.
The lack of an intuitive GUI and visualization \footnote{which are listed in the
non-features list on the web page:
\url{http://lammps.sandia.gov/non_features.html}} makes the workflow somewhat
complicated.

Atomify solves combines script editing, simulation,
visualization, and analysis in one single software application.
It uses LAMMPS as a library and synchronizes after each timestep,
copying the relevant information from LAMMPS to the graphics processing unit
(GPU) for visualization and data plotting.
The visualization is optimized for high performance by raytracing on billboards.
This allows visualization of millions of atoms with good framerate on desktop
computers. Atomify is designed with simplicity as main goal.

@@

A simplified version of Atomify is made available for mobile platforms (Android
and iOS).
This version of Atomify provides access to some of the examples and is intended
for exploration of molecular dynamics simulations by students.
It also works as a demonstration of the visualization capabilities of the
desktop version of the software.
The authors are optimistic that teachers can use the visualizations provided by
Atomify to explain microscopic interactions underlying concepts in topics such
as thermodynamics, geophysics, and biology.

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{final_billboard.png}
	\caption{Oversikt over GUI}
	\label{fig:gui}
\end{figure}

\section{Features}

@@

Atomify provides a lightweight script editor with syntax highlighting and line
numbers.
% TODO add built-in help and autocompletion to Atomify and mention here
Multiple files can be open at the same time and persist across working sessions.
The script can easily be restarted by clicking the button in the GUI or using
the default shortcut (Ctrl+R in Windows and Linux or Cmd+R in macOS).
Any changes made to the script are immediately visualized in the simulation.

Certain Atomify-commands can be added to the comments of a regular LAMMPS script
to modify properties of the GUI.
For instance, the color and size of the atoms can be set based on the type by
the command \texttt{\#/atom}.
To set the color to white and radius to 2.0 for particle type 1,
the user can apply the command \texttt{\#/atom 1 2.0 \#ffffff}.
The initial camera position can also be set similarily.
For a full list of the available commands, see the Atomify documentation.

@@

\bibliographystyle{plain}
\bibliography{Remote}

>>>>

\documentclass[aps,pre,twocolumn,letterpaper,floatfix,nofootinbib]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\usepackage{todo}
\usepackage{menukeys}

\begin{document}

@@

\section{Introduction}
%
Molecular dynamics (MD) is a common technique for simulating the motion of atoms, molecules and larger particles in a broad range of fields such as physics, chemistry and biology \cite{flerefelt}.
% TODO does MD also include larger particles?
% TODO which fields?
Over the years, increasingly sophisticated methods have been developed.
% TODO what methods?
With the large number of force fields, advanced techniques and GPU acceleration,
implementing a good molecular dynamics simulation is a challenging task.
There are many software packages for molecular modelling
available\footnote{\url{https://en.wikipedia.org/wiki/Comparison_of_software_for_molecular_mechanics_modeling} compares different softwares for molecular modeling.}.
Some packages, such as LAMMPS\cite{Plimpton1995Fast}, GROMACS\cite{Pronk2013}
and NAMD\cite{Phillips2005Scalable} have been developed for more than two decades 
and contain more than one million lines of code each.

The work in this paper is focused on LAMMPS.
Although the software is well-documented and enables the user to perform
advanced simulations, it can be demanding to learn how to use it well.
A LAMMPS simulation is performed by a series of commands which are executed one by
one.
Physical quantities such as energy, temperature and stress can be measured by
adding \textit{computes}.
The time evolution of these quantities can in turn be saved to file for plotting
and further analysis.
Particle trajectories can be saved to a file and visualized by tools such as
VMD\cite{Humphrey1996Vmd} or OVITO\cite{Stukowski2009Visualization}.
The lack of an intuitive GUI and visualization\footnote{GUI is listed as a non-feature on the web page:
\url{http://lammps.sandia.gov/non_features.html}.} makes the workflow somewhat
complicated.

Atomify solves combines script editing, simulation,
visualization, and analysis in one single software application.
It uses LAMMPS as a library and synchronizes after each timestep,
copying the relevant information from LAMMPS to the graphics processing unit
(GPU) for visualization and data plotting.
The visualization is optimized for high performance by raytracing billboards.
This allows visualization of millions of atoms with good framerate on desktop
computers. Atomify is designed with simplicity as main goal.

@@

A simplified version of Atomify is made available for mobile platforms (Android
and iOS).
This version of Atomify provides access to some of the examples and is intended
for exploration of molecular dynamics simulations by students.
It also works as a demonstration of the visualization capabilities of the
desktop version of the software.
The authors are optimistic that teachers can use the visualizations provided by
Atomify to explain microscopic interactions underlying concepts in topics such
as thermodynamics, geophysics, and biology.
\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{gui.png}
	\caption{Overview of the Graphical User Interface (GUI) in Atomify.}
	\label{fig:gui}
\end{figure*}
\section{Features}

@@

Atomify provides a lightweight script editor with syntax highlighting and line
numbers.
% TODO add built-in help and autocompletion to Atomify and mention here
Multiple files can be open at the same time and persist across working sessions.
The script can easily be restarted by clicking the button in the GUI or using
the default shortcut (\keys{Ctrl+R} in Windows and Linux or \keys{\cmd R} in macOS).
This enables you to quickly do changes in the script and see the changes immediately.

Certain Atomify-commands can be added to the comments of a regular LAMMPS script
to modify properties of the GUI. All these start with \keys{\texttt{\#}} so that
LAMMPS ignores them as they are interpreted as comments.
For instance, the color and size of the atoms can be set based on the type by
the command \keys{\texttt{\#/atom}}.
To set the color to white and radius to 2.0 for particle type 1,
the user can apply the command \keys{\texttt{\#/atom 1 2.0 \#ffffff}}. Certain built-in atom
types have default values and can be accessed with \keys{\texttt{\#/atom 1 oxygen}} for simplicity.
The initial camera position can also be set similarily.
For a full list of the available commands, see the Atomify documentation.

@@

% \bibliographystyle{plain}
\bibliography{Remote}
